GrapeCPNet: A Self-supervised Point Cloud Completion Network for 3D Phenotyping of Grape Bunches

### 2025-05-02 response2reviewer.tex

\usepackage{natbib}

% 关联主文档用来索引引用
\usepackage{xr}
\externaldocument{main}

\usepackage[draft,commandnameprefix=ifneeded]{changes}
\definechangesauthor[color=red]{R1}
\definechangesauthor[color=blue]{R2}

\rfoot{Page \thepage}

\lettertitle{The Response to Reviewers}
\title{GrapeCPNet: A Self-supervised Point Cloud Completion Network for 3D Phenotyping of Grape Bunches}
\author{
    Wenli Zhang \textsuperscript{a,*},
    Chao Zheng \textsuperscript{a},
    Chenhuizi Wang \textsuperscript{a},
    Pieter M. Blok \textsuperscript{b},
    Haozhou Wang \textsuperscript{b},
    Wei Guo \textsuperscript{b,*}
}
\journal{Computers and Electronics in Agriculture}
\doi{COMPAG-D-25-00311R1}

@@

\corresponding{Corresponding author: zhangwenli@bjut.edu.cn; guowei@g.ecc.u-tokyo.ac.jp }
\email{
    \textsuperscript{a} Information Department, Beijing University of Technology, Beijing, China \\
    \textsuperscript{b} Graduate School of Agricultural and Life Sciences, The University of Tokyo, Tokyo, Japan
}

\begin{document}

\begin{center}
    \maketitle
\end{center}

%%%%%%%%%%%%%%%%
% cover letter %
%%%%%%%%%%%%%%%%
\thedate

Dear Editor:

Thank you for giving us the opportunity to submit a revised draft of the manuscript ``\thetitle'' for publication in the Journal of \thejournal. We appreciate the time and effort that editors and the reviewers dedicated to providing feedback on our manuscript and are grateful for the insightful comments and valuable improvements to our paper. We have incorporated most of the suggestions made by the reviewers. Those changes are highlighted within the manuscript. Please see below, for a point-by-point response to the reviewers' comments and concerns. All page numbers refer to the revised manuscript file with tracked changes.

Thank your for your consideration. I am looking forward to hearing from you soon.

Sincerely,

Wei Guo\\
Associate professor\\
Graduate School of Agricultural and Life Science\\
The University of Tokyo, Tokyo, Japan\\
Email: guowei@g.ecc.u-tokyo.ac.jp

\vfill
\textbf{Note:} To enhance the legibility of this response letter, all the editor's and reviewers' comments are typeset in boxes. Rephrased or added sentences are typeset in color. The respective parts in the manuscript are highlighted to indicate changes.

%====================
% Response to Editor
%====================
\editor

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{reviewercomment}
    Overall, the paper is well structured with a clear goal,  and the authors provide insight to justify their design choices. However, I believe the paper will benefit from  more experiments, particularly on the instance segmentation and  completion parts.
\end{reviewercomment}

%============
% Reviewer 1
%============
\reviewer
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{reviewercomment}
    Overall, the paper is well structured with a clear goal,  and the authors provide insight to justify their design choices. However, I believe the paper will benefit from  more experiments, particularly on the instance segmentation and  completion parts.
\end{reviewercomment}

### 2025-05-02 response2reviewer.tex

Wei Guo\\
Associate professor\\
Graduate School of Agricultural and Life Science\\
The University of Tokyo, Tokyo, Japan\\
Email: guowei@g.ecc.u-tokyo.ac.jp

\vfill
\textbf{Note:} To enhance the legibility of this response letter, all the editor's and reviewers' comments are typeset in boxes. Rephrased or added sentences are typeset in color. The respective parts in the manuscript are highlighted to indicate changes.

%====================
% Response to Editor
%====================
\editor

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{reviewercomment}
    Thank you for submitting your manuscript to Computers and Electronics in Agriculture. 
    I have received comments from reviewers on your manuscript. 
    Your paper should become acceptable for publication pending suitable moderate revision and modification of the article in light of the appended reviewer comments.

    When resubmitting your manuscript, please carefully consider all issues mentioned in the reviewers' comments, outline every change made point by point, and provide suitable rebuttals for any comments not addressed.
\end{reviewercomment}

\response{
    We sincerely appreciate the positive feedback from the editor and reviewers regarding our manuscript. 
    We are grateful for the opportunity to revise our work based on the valuable comments received.

    In this revision, we have:

    1. Carefully addressed all reviewer comments point-by-point;\\
    2. Provided detailed responses explaining each modification;\\
    3. Made all necessary revisions to improve manuscript clarity and technical accuracy;\\
    4. Included tracked changes in the revised manuscript for transparent review.

    We believe these revisions have significantly strengthened the paper, and we appreciate the editor's and reviewers' time and constructive suggestions. 
    Please find our point-by-point responses to each reviewer comment in the following sections.
}

@@

%============
% Reviewer 1
%============
\reviewer
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{reviewercomment}
    Overall, the paper is well structured with a clear goal, and the authors provide insight to justify their design choices. 
    However, I believe the paper will benefit from  more experiments, particularly on the instance segmentation and completion parts.
\end{reviewercomment}

\response{
    We sincerely appreciate the reviewer's positive assessment of our study and manuscript. 
    Regarding the suggestion for additional experiments, we agree that further validation would strengthen the work.

    In response, we have:
}

\todoblock{
    Summary of added experiment.
}

@@

\response{
    We appreciate the reviewer's comment. The term "instance-level" refers to the individual segmentation of each grape berry and stem in the point cloud, where each instance is distinctly labeled and separated from others. We have clarified this definition in the revised manuscript.
}

\manuscript{
    In order to train and validate SoftGroup, each berry and stem in the bunch point cloud obtained in Section~\ref{sec:212} was \added[id=R2]{individually} labeled \replaced[id=R2]{(instance-level segmentation) as shown in}{at the instance-level} Fig.~\ref{fig:raw11}.
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{reviewercomment}
    Which platform, software, or tool was used for labeling?
\end{reviewercomment}

\response{
    We thank the reviewer for raising this important methodological detail. 
    The 3D point clouds were labeled using CloudCompare (v2.12.4), an open-source software for 3D point cloud and mesh processing.
}

\todoblock{
    Confirm which tool for labeling
}

@@

    $\hat{P}_C \gets \text{Normalize}(\text{Translate}(P_C))$ \tcp{center to $(0,0,0)$, range to $[-0.5m, 0.5m]$}
    \For{$i = 0 \to n$}{
        set $m = \text{random}\{1,2,3,4,5\}$ \tcp{\added[id=R2]{number of removal sphere regions}}
        $P_{o_i} \gets copy(\hat{P}_C)$ \tcp{initialize one output}
        \For{$j = 0 \to m$}{
            set $O = \{(x_o, y_o, z_o) \mid x, y, z \in \text{random}[-0.75, 0.75]\}$ \tcp{\added[id=R2]{removal sphere} center} 
            set $R = \text{random}[0.25, 0.75]$ \tcp{\added[id=R2]{removal sphere} radius}
            set $M_S \gets \text{sphere}(O, R)$ \tcp{generate \added[id=R2]{removal sphere region}}
            $P_{o_i} = P_{o_i} - P_{o_i} \cap M_S$ \tcp{remove overlap \added[id=R2]{within sphere region}}
        }
        $P_{\text{output}_i} \gets P_{o_i}$ \tcp{add to output set}
    }
\end{algorithm}

\begin{figure}[H]
    \centering
    \includegraphics[width=1\textwidth]{figures/Figure9.pdf}
    \caption{\replaced[id=R2]{Workflow of automatic paired training dataset generation}{Example of dataset composition} for the \deleted[id=R2]{self-supervised training-based method of the} berry point cloud completion network.}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{reviewercomment}
    Table 3 results: Was the same dataset used to assess the other algorithms, or were those metrics taken from their original studies?
\end{reviewercomment}

\response{
    We appreciate the reviewer's attention to experimental rigor. 
    To ensure fair comparison, all the comparing methods were conducted on our grape dataset under the same conditions.
    We have added this clarification to the results part.
}

\manuscript{
    GrapeCPNet achieves the best results under the self-supervised training method in both the training phase and the application phase, better than the comparative algorithms \added[id=R2]{under the same condition of our grape dataset}. 
}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{reviewercomment}
    Occlusion threshold: Is there a berry-occlusion threshold? 
    For example, if less than 50\% of a berry is visible, is it included? How are barely visible berries handled?
\end{reviewercomment}

\response{
    We thank the reviewer for this insightful question regarding occlusion handling. 
    Our approach was designed specifically to address challenging occlusion scenarios.

    We intentionally avoided setting visibility thresholds (e.g., 50\% cutoff) to maintain the applicable of proposed method.
    As shown in Fig.~\ref{fig:raw12}b, after removing 4 sphere regions, the incomplete berry for network training can achieved nearly around 75\% occlusion. 
    Thus, the GrapeCPNet is able to learn serever occlusions for barely visible berries.

    However, we must also admit that when severely obscured, only partial information can be seen. 
    The more obscured it is, the less information it provides. 
    Therefore, errors inevitably increase when the berry is not in a regular shape.
    As stated in the discussion part, we planed to taking prior knowledge of berry geometry to further improve the quality of completion.
}

\manuscript{
    We found that GrapeCPNet performed a little bit worse in the case of strange berry shapes, such as drop-shaped \added[id=R2]{berries, particularly when berries were sereverly occluded}.
    We subsequently consider utilizing the prior knowledge of berry geometry to achieve more detailed morphological feature extraction and representation to improve the quality of completion. 
}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{reviewercomment}
    Limitations \& future directions: How will the  model perform on wine grapes. Wine grapes accounts for ~60\% of US grape  production; clusters are denser, and berries are smaller than in table grapes.
\end{reviewercomment}

\response{
    We appreciate the reviewer's insightful comment regarding wine grapes, which indeed represent a major segment of grape production. 
    In our current study, we just focused on table grape varieties commonly found in Chinese markets.
    Although they covered a wide range of cluster densities (from sparse to very dense) and berry sizes (20-35 cm axis length, as shown in Table~\ref{tbl:1} and Figure~\ref{fig:raw17}), 
    we acknowledge that wine grapes typically have smaller berry sizes lower than 13-15 mm \citep{manso_wine_2021,melo_berry_2015}, these small sizes were not included in our dataset.

    From a deep learning perspective, our model architecture should be capable of handling smaller wine grapes if provided with appropriate training data, as the fundamental detection principles remain similar regardless of grape variety. 
    Nevertheless, we agree that future work should specifically address wine grape detection through collaboration with U.S. researchers to: 
    (1) collect wine grape datasets, and 
    (2) evaluate whether models trained on table grapes can generalize well to the smaller wine grape berries without additional training.

    We have added the previous points to the discussion parts.
}

\manuscript{
    \added[id=R2]{
        This study has another limitation: it focuses only on four table grapes, which are currently common and available in Chinese markets.
        Although we examined a wide range of cluster densities (from sparse to very dense) and berry sizes (20-35 mm axis length, as shown in Table~\ref{tbl:1} and Figure~\ref{fig:raw17}),
        the study did not include wine grapes, which account for over 60\% of U.S. grape production and typically have smaller berries and denser clusters.
        As \citet{manso_wine_2021,melo_berry_2015} report, wine grape berries are often smaller than 13-15 mm, a size range not covered in our dataset.
        From a deep learning perspective, our model architecture should be able to detect smaller wine grapes with appropriate training data, as the basic detection principles are similar across grape varieties.
        Future work should involve collaboration with U.S. researchers to collect wine grape datasets and assess whether models trained on table grapes can generalize well to smaller wine grape berries without additional training.
    }

    \replaced[id=R2]{Additionally, this study only}{Another limitation of this study is that it} focused on a single berry bunch in a controlled environment.
}

% \phantom{\cite{}} % 添加一个不可见的伪引用, 不会显示但能满足BibTeX要求避免无引用时报错
\bibliographystyle{elsarticle-harv}
% \renewcommand{\bibsection}{} % 隐藏reference标题，如果需要则注释掉这一行
\bibliography{references}

### 2025-05-02 main.tex
response to R2

The berry point cloud completion experiments are divided into two parts according to the different phases of the network: 
the results of the self-supervised training (Test) and the results of the practical application (Inference).
As shown in Table~\ref{tbl:4}, 
GrapeCPNet achieves the best results under the self-supervised training method in both the training phase and the application phase, better than the comparative algorithms. Fig.~\ref{fig:raw14} visualizes the shape completion result of GrapeCPNet on the incomplete berry segmented from the bunch in the application phase.

@@

For berry completion using deep learning, the most challenging and time-consuming task is preparing the paired training datasets of complete and incomplete point clouds.
In our previous work on potato tuber completion networks \citep{blok_highthroughput_2025}, a distinguishable pin was used as a reference point to align the complete tuber scans obtained via photogrammetry with the incomplete tuber scans captured using an RGB-D camera.
A pin-guided point cloud matching algorithm was then applied to calculate the transformation matrix for pairing the datasets.
Potato tubers, being single objects with clear textures and shape variation, presented fewer challenges when aligning data collected by two different types of sensors.
However, grape berries pose a significantly more difficult problem due to their highly rounded shapes, weak texture features, and strong reflectance, which complicates the creation of paired data.
As a result, the previous pin-guided point cloud pairing method used for potato tuber datasets is not effective for grape berries.
Thus, we simulated random sphere cutting on the complete berry point clouds to generate thousands of paired incomplete point clouds, as described in Subsection~\ref{sec:berrycut}.
As demonstrated in Subsection~\ref{sec:32}, these automatically generated paired training datasets were successfully used to train several completion models (Table~\ref{tbl:4}).
We have found that some researchers \citep{pan_panoptic_2023,magistri_efficient_2024} have improved the completion effect by encoding a prior knowledge of the geometry of the object into the weights of the neural network. 
Considering the agricultural plant completion task, our proposed GrapeCPNet, which combines centroid-based sparse completion with diffusion-based dense completion, outperformed other networks in terms of performance.
We found that GrapeCPNet performed a little bit worse in the case of strange berry shapes, such as drop-shaped.
We subsequently consider utilizing the prior knowledge of berry geometry to achieve more detailed morphological feature extraction and representation to improve the quality of completion.

% challenges, limitations and future works.
Another limitation of this study is that it focused on a single berry bunch in a controlled environment.
The noise in the data is also related to the effectiveness of the method proposed in this paper.
For practical applications in vineyards, collecting the same point cloud quality as that achieved in controlled indoor environments can be challenging, and multi-view data collection may not be as efficient.
How to effectively implement the method in practical application scenarios is another issue to be considered in future research.
To increasing data collection efficiency, devices such as RGB-D cameras \citep{blok_highthroughput_2025}, fisheye cameras \citep{tamura_unsupervised_2025}, or even 360 cameras can be used.
Moreover, as in previous studies \citep{rose_automated_2016, luo_3d_2023}, investigating methods to segment individual grape bunches directly from vineyard scans using deep learning techniques should be an area of future exploration.
Additionally, \citet[Fig.~9]{scholer_automated_2015} successfully demonstrated the construction of virtual 3D models of grape bunches, including the hidden branching architecture. 
By collecting time-series growth data, the branching structure could potentially be captured during early stages of thinning, as suggested by \citep{du_instance_2023}.
Such data could enable studies on growth prediction for optimizing harvesting times and assessments of branching structure deviations within crowded berry bunches.
Finally, applying the latest procedural computer graphics (CG) modeling techniques or functional-structural plant models represents another promising approach to advancing practical applications in this field.

>>>>

The berry point cloud completion experiments are divided into two parts according to the different phases of the network: 
the results of the self-supervised training (Test) and the results of the practical application (Inference).
As shown in Table~\ref{tbl:4}, 
GrapeCPNet achieves the best results under the self-supervised training method in both the training phase and the application phase, better than the comparative algorithms \added[id=R2]{under the same condition of our grape dataset}. 
Fig.~\ref{fig:raw14} visualizes the shape completion result of GrapeCPNet on the incomplete berry segmented from the bunch in the application phase.

@@

For berry completion using deep learning, the most challenging and time-consuming task is preparing the paired training datasets of complete and incomplete point clouds.
In our previous work on potato tuber completion networks \citep{blok_highthroughput_2025}, a distinguishable pin was used as a reference point to align the complete tuber scans obtained via photogrammetry with the incomplete tuber scans captured using an RGB-D camera.
A pin-guided point cloud matching algorithm was then applied to calculate the transformation matrix for pairing the datasets.
Potato tubers, being single objects with clear textures and shape variation, presented fewer challenges when aligning data collected by two different types of sensors.
However, grape berries pose a significantly more difficult problem due to their highly rounded shapes, weak texture features, and strong reflectance, which complicates the creation of paired data.
As a result, the previous pin-guided point cloud pairing method used for potato tuber datasets is not effective for grape berries.
Thus, we simulated random sphere cutting on the complete berry point clouds to generate thousands of paired incomplete point clouds, as described in Subsection~\ref{sec:berrycut}.
As demonstrated in Subsection~\ref{sec:32}, these automatically generated paired training datasets were successfully used to train several completion models (Table~\ref{tbl:4}).
We have found that some researchers \citep{pan_panoptic_2023,magistri_efficient_2024} have improved the completion effect by encoding a prior knowledge of the geometry of the object into the weights of the neural network. 
Considering the agricultural plant completion task, our proposed GrapeCPNet, which combines centroid-based sparse completion with diffusion-based dense completion, outperformed other networks in terms of performance.
We found that GrapeCPNet performed a little bit worse in the case of strange berry shapes, such as drop-shaped \added[id=R2]{berries, particularly when berries were sereverly occluded}.
We subsequently consider utilizing the prior knowledge of berry geometry to achieve more detailed morphological feature extraction and representation to improve the quality of completion.

% challenges, limitations and future works.
\added[id=R2]{
    This study has another limitation: it focuses only on four table grapes, which are currently common and available in Chinese markets.
    Although we examined a wide range of cluster densities (from sparse to very dense) and berry sizes (20-35 mm axis length, as shown in Table~\ref{tbl:1} and Figure~\ref{fig:raw17}),
    the study did not include wine grapes, which account for over 60\% of U.S. grape production and typically have smaller berries and denser clusters.
    As \citet{manso_wine_2021,melo_berry_2015} report, wine grape berries are often smaller than 13-15 mm, a size range not covered in our dataset.
    From a deep learning perspective, our model architecture should be able to detect smaller wine grapes with appropriate training data, as the basic detection principles are similar across grape varieties.
    Future work should involve collaboration with U.S. researchers to collect wine grape datasets and assess whether models trained on table grapes can generalize well to smaller wine grape berries without additional training.
}

\replaced[id=R2]{Additionally, this study only}{Another limitation of this study is that it} focused on a single berry bunch in a controlled environment.
The noise in the data is also related to the effectiveness of the method proposed in this paper.
For practical applications in vineyards, collecting the same point cloud quality as that achieved in controlled indoor environments can be challenging, and multi-view data collection may not be as efficient.
How to effectively implement the method in practical application scenarios is another issue to be considered in future research.
To increasing data collection efficiency, devices such as RGB-D cameras \citep{blok_highthroughput_2025}, fisheye cameras \citep{tamura_unsupervised_2025}, or even 360 cameras can be used.
Moreover, as in previous studies \citep{rose_automated_2016, luo_3d_2023}, investigating methods to segment individual grape bunches directly from vineyard scans using deep learning techniques should be an area of future exploration.
Additionally, \citet[Fig.~9]{scholer_automated_2015} successfully demonstrated the construction of virtual 3D models of grape bunches, including the hidden branching architecture. 
By collecting time-series growth data, the branching structure could potentially be captured during early stages of thinning, as suggested by \citep{du_instance_2023}.
Such data could enable studies on growth prediction for optimizing harvesting times and assessments of branching structure deviations within crowded berry bunches.
Finally, applying the latest procedural computer graphics (CG) modeling techniques or functional-structural plant models represents another promising approach to advancing practical applications in this field.
